\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-06-20 paper/subdocuments/results.Rnw
integrated HABE comparison

<<different_ginis>>=
df$G_reink[df$G_reink==1]<-NA
df$G_taxed[df$G_taxed==1]<-NA
library(reshape)
dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
names(dflong)[3]<-"Type"
names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("taxable income","income before deductions","taxable income after tax")
ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Type))+
  scale_y_continuous(limits=c(0.18,0.4))+
  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
  geom_point(aes(shape=Type),size=3)+
  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Type))+
  theme_bw()+
  xlab("tax period")
#ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
@

% Nicht eher nur CH-Plot?

@@

>>>>

<<different_ginis>>=
df$G_reink[df$G_reink==1]<-NA
df$G_taxed[df$G_taxed==1]<-NA
library(reshape)
number_ticks <- function(n) {function(limits) pretty(limits, n)}
dflong <- melt(df[,c("G_steink","G_reink","G_taxed","kanton","steuerperiode")], id=c("kanton","steuerperiode"))
names(dflong)[3]<-"Type"
names(dflong)[4]<-"Gini"
levels(dflong$Type)<-c("taxable income","income before deductions","taxable income after tax")
ggplot(dflong[dflong$kanton=="CH",], aes(x=steuerperiode,y=Gini,shape=Type))+
  scale_y_continuous(limits=c(0.18,0.4))+
  scale_x_continuous(limits=c(1982,2012),breaks=number_ticks(10)) +
  geom_point(aes(shape=Type),size=3)+
  geom_line(data=dflong[dflong$kanton=="CH",],aes(linetype=Type))+
  theme_bw()+
  xlab("tax period")
#ggplot(dflong, aes(x=steuerperiode,y=Gini,colour=Type))+geom_line()+facet_wrap(~kanton)+theme_bw()+xlab("tax period")
@

\subsection{Statistical units}

<<setup_data>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
# Ohne Nuller
bd <- read.csv("../../data/data_Schweiz.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd$Jahr <- as.numeric(substr(bd$Veranlagungsperiode,1,4))
bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] <- bd$Jahr[nchar(as.character(bd$Veranlagungsperiode))>4] - 2
bd<-bd %.% 
  filter(Einheit=="Total")
bd<-subset(bd,Jahr<1995 | Jahr >2002)
bd.long<-reshape(bd,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income (only taxed)","Equivalised net income (only taxed)"),
direction ="long")
# mit Nuller
bd.0 <- read.csv("../../data/data_Schweiz_mitNull.csv", header=TRUE,sep=";")
#Veranlagungsperiode -> Jahre
bd.0$Jahr <- as.numeric(substr(bd.0$Veranlagungsperiode,1,4))
bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] <- bd.0$Jahr[nchar(as.character(bd.0$Veranlagungsperiode))>4] - 2
bd.0<-bd.0 %.% 
  filter(Einheit=="Total")
bd.0<-subset(bd.0,Jahr<1995 | Jahr >2002)
bd.0.long<-reshape(bd.0,
varying = c("gini_reink","gini_reinka"),
v.names = "Gini",
timevar = "inc_def",
times = c("Net income","Equivalised net income"),
direction ="long")
bdlong<-rbind(bd.0.long,bd.long)
@

<<With and without equivalence scale>>=
library(ggplot2)
number_ticks <- function(n) {function(limits) pretty(limits, n)}
ggplot(bdlong, aes(x=Jahr,y=Gini,shape=inc_def))+
  scale_y_continuous(limits=c(0.18,0.6))+
  scale_x_continuous(limits=c(1940,2012),breaks=number_ticks(10)) +
  geom_point(aes(shape=inc_def),size=3)+
  geom_line(data=bdlong,aes(linetype=inc_def))+
  theme_bw()+
  xlab("tax period")
@

@@

<<setup_data>>=
library(dplyr, quietly=TRUE, warn.conflicts = FALSE)
library(ggplot2)
library(reldist)
library(foreign)
# Get Brülhartdata for 2010
estv2010 <- read.csv("data/estv_normal_sonder_mitNuller_stE_2010.csv",sep=";")
###########################################
# Select percentiles for all cases (total)
estv2010.total<-estv2010[c(18:38),c(1,4)]
##
# Simulate individual data points
# cumulative probabilites
cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
# probabilities
prob<-c(cum.p[1],diff(cum.p), .01)
# extreme values beyond x (to sample)
#freq<-max(normal$anz_pflichtige)
freq<-3000
init<-0
fin<-(abs(max(estv2010.total$Total,na.rm=TRUE))+5)
# generate the sequence to take pairs from
ival<-c(init,estv2010.total$Total,fin)
len<-10000 # sequence of each pair
s<-sapply(2:length(ival),function(i){
  seq(ival[i-1],ival[i],length.out=len)
})
s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
estv2010.total.ind<-data.frame(s)
names(estv2010.total.ind)[1]=c("inc")
estv2010.total.ind$inc<-estv2010.total.ind$inc/1000

# HABE - 2010 all cases

habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
habe10<-habe0911 %.% 
  filter(Jahr08==2010)
habe10<-data.frame(habe10$VerfuegbaresEinkommen08)
names(habe10)<-"inc_habe"
habe10$inc_habe[habe10$inc_habe<0]<-0 # One values is -20'000
habe10$inc_habe<-habe10$inc_habe*12/1000 # transform from monthly income to income per year

# Probability density Function

density1 <- density(habe10$inc_habe)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "Einkommen", ylab = "density",
     axes = FALSE,
     xlim = c(-70, 4600),
     ylim=c(0,1.103e-02))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(2500,2500),y=c(1.103e-02,1.103e-02))
legend(fig1legend,lty=1:2,cex=0.5, bty="n",
       legend=c("HBS","FTA"))
density2 <- density(estv2010.total.ind$inc)
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

# Relative Distribution

par(mfrow=c(1,3))
g10 <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
               smooth=0.4, ci=FALSE,
               yolabs=seq(-1,3,by=0.5),
               ylim=c(0.5,5.0),
               bar=TRUE, quiet=FALSE,
               xlab="Proportion of HBS Distribution")
title(main=paste("Overall realtive density"))
abline(h=1,lty=2)
g1A <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
               show="effect",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,5.0), ylab="",
               smooth=0.4, ci=FALSE,
               yolabs=seq(-1,3,by=0.5),
               xlab="Proportion of HBS Distribution")
title(main=paste("Effect of Median shift"))
abline(h=1,lty=2)
gA0 <- reldist(y=estv2010.total.ind$inc, yo=habe10$inc_habe,
               smooth=0.4, ci=FALSE,
               show="residual",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,5.0), ylab="",
               yolabs=seq(-1,3,by=0.5),
               xlab="Proportion of HBS Distribution")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)
par(mfrow=c(1,1))

########################
# Separate comparison for unmarried

# Select married - ESTV

estv2010.married<-estv2010[c(18:38),c(1,2)]
names(estv2010.married)[2]=c("married")
##
# Simulate individual data points
# Percentile-Plot (descriptive  purpose)
# cumulative probabilites
cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
# probabilities
prob<-c(cum.p[1],diff(cum.p), .01)
# extreme values beyond x (to sample)
#freq<-max(normal$anz_pflichtige)
freq<-3000
init<-0
fin<-(abs(max(estv2010.married$married,na.rm=TRUE))+5)
# generate the sequence to take pairs from
ival<-c(init,estv2010.married$married,fin)
len<-10000 # sequence of each pair
s<-sapply(2:length(ival),function(i){
  seq(ival[i-1],ival[i],length.out=len)
})
s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
estv2010.married.ind<-data.frame(s)
names(estv2010.married.ind)[1]=c("inc")
estv2010.married.ind$inc<-estv2010.married.ind$inc/1000

# Select married - HABE

habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
habe10<-habe0911 %.% 
  filter(Jahr08==2010)
habe10.married<-habe0911 %.% 
  filter(Familientyp08>130,Jahr08==2010)
habe10.married<-data.frame(habe10.married$VerfuegbaresEinkommen08)
names(habe10.married)<-"inc_habe"
habe10.married$inc_habe[habe10.married$inc_habe<0]<-0 # One values is -20'000
habe10.married$inc_habe<-habe10.married$inc_habe*12/1000 # transform from monthly income to income per year

# Probability Density Function

density1 <- density(habe10.married$inc_habe)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "Einkommen", ylab = "density",
     axes = FALSE,
     xlim = c(-257, 8000),
     ylim=c(0,0.0101390))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(2500,2500),y=c(0.0101390,0.0101390))
legend(fig1legend,lty=1:2,cex=0.5, bty="n",
       legend=c("HBS","FTA"))
density2 <- density(estv2010.married.ind$inc)
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

# Relative Distribution

par(mfrow=c(1,3))
g10 <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
               smooth=0.4, ci=FALSE,
               yolabs=seq(-1,3,by=0.5),
               ylim=c(0.5,5.0),
               bar=TRUE, quiet=FALSE,
               xlab="Proportion of HBS Distribution")
title(main=paste("Overall realtive density"))
abline(h=1,lty=2)
g1A <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
               show="effect",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,5.0), ylab="",
               smooth=0.4, ci=FALSE,
               yolabs=seq(-1,3,by=0.5),
               xlab="Proportion of HBS Distribution")
title(main=paste("Effect of Median shift"))
abline(h=1,lty=2)
gA0 <- reldist(y=estv2010.married.ind$inc, yo=habe10.married$inc_habe,
               smooth=0.4, ci=FALSE,
               show="residual",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,5.0), ylab="",
               yolabs=seq(-1,3,by=0.5),
               xlab="Proportion of HBS Distribution")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)
par(mfrow=c(1,1))

#################
# Separate comparison for unmaried

##
# Unmarried ESTV

estv2010.unmarried<-estv2010[c(18:38),c(1,3)]
names(estv2010.unmarried)[2]=c("unmarried")
##
# Simulate individual data points
# Percentile-Plot (descriptive  purpose)
# cumulative probabilites
cum.p<-c(1,5,10,20,25,30,40,50,60,70,75,80,90,95,96,97,98,99,99.5,99.9,99.99)/100
# probabilities
prob<-c(cum.p[1],diff(cum.p), .01)
# extreme values beyond x (to sample)
#freq<-max(normal$anz_pflichtige)
freq<-3000
init<-0
fin<-(abs(max(estv2010.unmarried$unmarried,na.rm=TRUE))+5)
# generate the sequence to take pairs from
ival<-c(init,estv2010.unmarried$unmarried,fin)
len<-10000 # sequence of each pair
s<-sapply(2:length(ival),function(i){
  seq(ival[i-1],ival[i],length.out=len)
})
s<-sample(s,freq,prob=rep(prob, each=len),replace=T)
estv2010.unmarried.ind<-data.frame(s)
names(estv2010.unmarried.ind)[1]=c("inc")
estv2010.unmarried.ind$inc<-estv2010.unmarried.ind$inc/1000

# Select unmarried - HABE

habe0911<-read.table("P:/WGS/FBS/ISS/Projekte laufend/SNF Ungleichheit/Datengrundlagen/HABE/2009 bis 2011/HABE091011_Standard_130717UOe.txt", header=TRUE)
habe10<-habe0911 %.% 
  filter(Jahr08==2010)
habe10.unmarried<-habe0911 %.% 
  filter(Familientyp08<140,Jahr08==2010)
habe10.unmarried<-data.frame(habe10.unmarried$VerfuegbaresEinkommen08)
names(habe10.unmarried)<-"inc_habe"
habe10.unmarried$inc_habe[habe10.unmarried$inc_habe<0]<-0 # One values is -20'000
habe10.unmarried$inc_habe<-habe10.unmarried$inc_habe*12/1000 # transform from monthly income to income per year

# Probability Density Function

density1 <- density(habe10.unmarried$inc_habe)
plot(x = (density1$x), y = density1$y, type = "l",
     xlab = "Einkommen", ylab = "density",
     axes = FALSE,
     xlim = c(-50, 2500),
     ylim=c(0,1.764e-02))
title(main="FTA vs HBS income distribution",cex=0.6)
axis(side = 1)
axis(side = 2)
fig1legend <- list(x=c(1200,1200),y=c(1.764e-02,1.764e-02))
legend(fig1legend,lty=1:2,cex=0.5, bty="n",
       legend=c("HBS","FTA"))
density2 <- density(estv2010.unmarried.ind$inc)
lines(x = (density2$x), y = density2$y, type = "l",lty=2)

# Relative Distribution

par(mfrow=c(1,3))
g10 <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
               smooth=0.4, ci=FALSE,
               yolabs=seq(-1,3,by=0.5),
               ylim=c(0.5,5.0),
               bar=TRUE, quiet=FALSE,
               xlab="Proportion of HBS Distribution")
title(main=paste("Overall realtive density"))
abline(h=1,lty=2)
g1A <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
               show="effect",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,5.0), ylab="",
               smooth=0.4, ci=FALSE,
               yolabs=seq(-1,3,by=0.5),
               xlab="Proportion of HBS Distribution")
title(main=paste("Effect of Median shift"))
abline(h=1,lty=2)
gA0 <- reldist(y=estv2010.unmarried.ind$inc, yo=habe10.unmarried$inc_habe,
               smooth=0.4, ci=FALSE,
               show="residual",
               bar=TRUE, quiet=FALSE,
               ylim=c(0.5,5.0), ylab="",
               yolabs=seq(-1,3,by=0.5),
               xlab="Proportion of HBS Distribution")
title(main=paste("Effect of different shape"))
abline(h=1,lty=2)
par(mfrow=c(1,1))
@
